---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r knitropts, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/readme-",
  fig.height = 4.5,
  fig.align = "center"
)
```

# ggopenair

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/ggopenair)](https://CRAN.R-project.org/package=ggopenair)
[![R-CMD-check](https://github.com/jack-davison/ggopenair/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jack-davison/ggopenair/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of `ggopenair` is to bridge the gap between the data analysis functions of `openair` and the flexibility of `ggplot2`. One of the key strengths of `ggplot2` over `lattice` is that it allows plots to be manipulated *after* they are created, which allows for a much more flexible plotting experience. `ggplot2` is also being actively developed and extended, which `ggopenair` can take advantage of.

## Installation

You can install the development version of ggopenair from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("jack-davison/ggopenair")
```

## Preview

A more thorough overview of `{ggopenair}` can be found on the package website: <https://jack-davison.github.io/ggopenair/>.

```{r demo, fig.height=8, fig.width=10, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggopenair)
library(ggplot2)
library(patchwork)

plt <- gg_polar_plot(marylebone, "nox") + theme_polar() +
  scale_y_continuous(breaks = seq(0, 12, 4))

classic <- plt + scale_opencolours_c() + ggtitle("Classic")

sqrt_trans <- plt + scale_opencolours_c(trans = c("sqrt")) + 
  ggtitle("Scale Transform")

steps <- plt + ggtitle("Binned") + 
  scale_color_stepsn(colours = openair::openColours(), 
                     n.breaks = 10)

annotated <- classic + ggtitle("Annotated") +
  annotate_polar_wedge("S", "W", fill = "black") +
  annotate("label", x = 225, y = 8, fill = "white", color = "red", 
           label = "watch out\nover here!")

classic + sqrt_trans + steps + annotated
```

