---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r knitropts, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.height = 4.5, 
  fig.align = "center"
)
```

# ggopenair

<!-- badges: start -->
<!-- badges: end -->

The goal of `ggopenair` is to bridge the gap between the data analysis functions of `openair` and the flexibility of `ggplot2`. One of the key strengths of `ggplot2` over `lattice` is that it allows plots to be manipulated *after* they are created, which allows for a much more flexible plotting experience. `ggplot2` is also being actively developed and extended, which `ggopenair` can take advantage of.

## Installation

You can install the development version of ggopenair from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("jack-davison/ggopenair")
```

## Demonstration

### Introduction

An openair polar plot looks like this:

```{r legacy}
openair::polarPlot(ggopenair::marylebone)
```

To achieve the same result in `ggopenair` one would write:

```{r ggpolar}
library(ggopenair)
library(ggplot2)

gg_polar_plot(marylebone, "nox") +
  theme_polar() +
  scale_opencolours()
```

This is more long winded, but the flexibility allows users to customise their outputs very closely. For example:

### Custom Themes

Use `ggplot2::theme()` or any theme package. For example, we can use the style from the Wall Street Journal:

```{r themes}
gg_polar_plot(openair::mydata, "nox") +
  ggthemes::theme_wsj() +
  scale_color_gradientn(colours = rev(ggthemes::wsj_pal()(2))) +
  guides(color = guide_colorbar(barwidth = grid::unit(5, "cm")))
```

Or we could make our polar plot blend in with some other plots produced in Google Sheets:

```{r}
gg_polar_plot(openair::mydata, "nox") +
  ggthemes::theme_gdocs() +
  scale_color_gradientn(colours = ggthemes::gdocs_pal()(3))
```

### Scales

Use any `ggplot2` scale function to change how the plot behaves. For example, use `scale_color_binned()` to bin the colour bar.

```{r scales}
gg_polar_plot(openair::mydata, "nox") +
  theme_polar() +
  scale_color_steps(low = "black",
                             high = "goldenrod",
                             breaks = seq(0, 1000, 40))
```

Alternatively, one could use the "trans" argument to shift the colour scale. This could be particularly useful for `gg_polar_freq()`, which had its own "trans" argument in `openair`.

```{r trans, fig.width=12}
shift_axis <- function(trans) {
  gg_polar_freq(marylebone, "nox") +
    theme_polar() +
    scale_fill_gradientn(colours = c("darkgreen", "hotpink"),
                                   trans = trans) +
    labs(title = trans)
}

patchwork::wrap_plots(shift_axis("identity"),
                      shift_axis("sqrt"),
                      shift_axis("log10"))
```

### Annotations

Use `annotate()` to easily draw on your polar plots and to draw attention to certain aspects. In-built annotation functions make it easy to, for example, draw a highlighting wedge or direct axis labels.

```{r annotations}
gg_polar_plot(openair::mydata, "nox") +
  theme_polar() +
  scale_opencolours("inferno") +
  annotate_polar_wedge("S", "W") +
  annotate_polar_axis(seq(5, 20, 5), color = "grey50")
```

Calendar plots also possess their own dedicated annotation functions.

```{r calendar, fig.width=10}
marylebone |>
  openair::selectByDate(year = 2019) |>
  gg_calendar("nox") +
  annotate_calendar_text("date") +
  annotate_calendar_wd(colour = "black") +
  scale_opencolours()
```

### Extension 

Use any of the `ggplot2` extension packages out there, such as `patchwork`. For example, a polar plot could be inserted into a time variation plot.

```{r patch, fig.height=6, fig.width=10, out.width="100%"}
library(patchwork)

polar <-
  gg_polar_plot(openair::mydata, "nox") + 
  theme_polar() + 
  theme(panel.border = element_rect(fill = NA, color = "black")) +
  scale_opencolours()

tv <- gg_timevariation(openair::mydata, "nox", return = "list")
tv <- purrr::map(tv, ~.x + theme_classic() + theme(legend.position = "none", panel.border = element_rect(fill = NA)))

tv$day_hour / (tv$month | tv$day | polar) +
  plot_layout(heights = c(.8, 1), guides = "collect")
```

