---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r knitropts, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.height = 4.5, 
  fig.align = "center"
)
```

# ggopenair

<!-- badges: start -->
<!-- badges: end -->

The goal of `ggopenair` is to bridge the gap between the data analysis functions of `openair` and the flexibility of `ggplot2`. One of the key strengths of `ggplot2` over `lattice` is that it allows plots to be manipulated *after* they are created, which allows for a much more flexible plotting experience. `ggplot2` is also being actively developed and extended, which `ggopenair` can take advantage of.

## Installation

You can install the development version of ggopenair from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("jack-davison/ggopenair")
```

## Example

An openair polar plot looks like this:

```{r legacy}
openair::polarPlot(openair::mydata)
```

To achieve the same result in `ggopenair` one would write:

```{r ggpolar}
library(ggopenair)

gg_polar_plot(openair::mydata, "nox") +
  theme_polar() +
  scale_opencolours()
```

This is more long winded, but the flexibility allows users to customise their outputs very closely. For example:

-  **Custom Themes:** Use `ggplot2::theme()` or any theme package. For example, we can use the style from the Wall Street Journal:

```{r themes}
gg_polar_plot(openair::mydata, "nox") +
  ggthemes::theme_wsj() +
  ggplot2::scale_color_gradientn(colours = rev(ggthemes::wsj_pal()(2))) +
  ggplot2::guides(color = ggplot2::guide_colorbar(barwidth = grid::unit(5, "cm")))
```

-  **Scales:** Use any `ggplot2` scale function to change how the plot behaves. For example, use `scale_color_binned()` to bin the colour bar.

```{r scales}
gg_polar_plot(openair::mydata, "nox") +
  theme_polar() +
  ggplot2::scale_color_binned(type = "viridis",
                              breaks = seq(0, 1000, 50))
```

-  **Annotations:** Use `annotate()` to easily draw on your polar plots to draw attention to certain aspects.

```{r annotations}
gg_polar_plot(openair::mydata, "nox") +
  theme_polar() +
  scale_opencolours("inferno") +
  annotate_polar_wedge("S", "W") +
  ggplot2::annotate(
    geom = "text",
    x = (270+180)/2, y = 12, angle = -45,
    label = "Looks high here!", 
    color = "white"
  )
```

-  **Extensions:** Use any of the `ggplot2` extension packages out there, such as `patchwork`. For example, a polar plot could be inserted into a time variation plot:

```{r patch, fig.height=6, fig.width=10, out.width="100%"}
library(patchwork)

polar <-
  gg_polar_plot(openair::mydata, "nox") + 
  theme_polar(guides = FALSE) + 
  ggplot2::theme(panel.border = ggplot2::element_rect(fill = NA, color = "black")) +
  scale_opencolours()

tv <- gg_timevariation(openair::mydata, "nox", return = "list")

(tv$day_hour + ggplot2::theme_bw() + ggplot2::theme(legend.position = "none")) /
  ((tv$month + ggplot2::theme_bw() + ggplot2::theme(legend.position = "none")) | 
     polar |
     (tv$day + ggplot2::theme_bw() + ggplot2::theme(legend.position = "none"))) +
  plot_layout(heights = c(.8, 1))

```

