---
title: "Polar Plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Polar Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

An `{openair}` polar plot looks like this:

```{r legacy}
openair::polarPlot(ggopenair::marylebone)
```

To achieve the same result in `{ggopenair}` one would write:

```{r ggpolar}
library(ggopenair)
library(ggplot2)

gg_polar_plot(marylebone, "nox") +
  theme_polar() +
  scale_opencolours_c()
```

This is more long winded, but the flexibility allows users to customise their outputs very closely. For example:

# Scales

Use any `ggplot2` scale function to change how the plot behaves. For example, use `scale_color_binned()` to bin the colour bar.

```{r scales}
gg_polar_plot(marylebone, "nox") +
  theme_polar() +
  scale_color_steps(
    low = "black",
    high = "goldenrod",
    breaks = seq(0, 1000, 40)
  )
```

Alternatively, one could use the "trans" argument to shift the colour scale. This could be particularly useful for `gg_polar_freq()`, which had its own "trans" argument in `openair`.

```{r trans, fig.width=12}
shift_axis <- function(trans) {
  gg_polar_freq(marylebone, "nox") +
    theme_polar() +
    scale_fill_gradientn(
      colours = c("darkgreen", "hotpink"),
      trans = trans
    ) +
    labs(title = trans)
}

patchwork::wrap_plots(
  shift_axis("identity"),
  shift_axis("sqrt"),
  shift_axis("log10")
)
```

# Annotations

Use `annotate()` to easily draw on your polar plots and to draw attention to certain aspects. In-built annotation functions make it easy to, for example, draw a highlighting wedge or direct axis labels.

```{r annotations}
gg_polar_plot(marylebone, "nox") +
  theme_polar() +
  scale_opencolours_c("inferno") +
  annotate_polar_wedge("S", "W") +
  annotate_polar_axis(seq(5, 15, 5), color = "grey50")
```

# Extensions

Use any of the `ggplot2` extension packages out there, such as `patchwork`. For example, a polar plot could be inserted into a time variation plot.

```{r patch, fig.height=6, fig.width=10, out.width="100%"}
library(patchwork)

polar <-
  gg_polar_plot(marylebone, "nox") +
  theme_polar() +
  theme(panel.border = element_rect(fill = NA, color = "black")) +
  scale_opencolours_c()

tv <- gg_timevariation(marylebone, "nox", return = "list")
tv <-
  purrr::map(
    tv,
    ~ .x + theme_classic() + theme(
      legend.position = "none",
      panel.border = element_rect(fill = NA)
    )
  )

tv$day_hour / (tv$month | tv$day | polar) +
  plot_layout(heights = c(.8, 1), guides = "collect")
```

# Other Polar Plots

Currently the main six `{openair}` polar plots have been implemented in `{ggopenair}`.

```{r allpolar, fig.width=7, fig.height=4}
polarplot <-
  gg_polar_plot(marylebone, "nox") + theme_polar() + scale_opencolours_c()
polarannulus <-
  gg_polar_annulus(marylebone, "nox") + theme_polar() + scale_opencolours_c()
polarfreq <-
  gg_polar_freq(marylebone, "nox") + theme_polar() + scale_opencolours_c(trans = "sqrt")
polarperc <-
  gg_polar_percentile(marylebone, "nox") + theme_polar() + scale_color_manual(values = openair::openColours(n = 5))
# pollrose <-
#   gg_pollutionrose(marylebone, "nox") + theme_polar(panel_ontop = FALSE) + scale_fill_manual(values = rev(openair::openColours(n = 7)))
# windrose <-
#   gg_windrose(marylebone) + theme_polar(panel_ontop = FALSE) + scale_fill_manual(values = rev(openair::openColours(n = 4)))

wrap_plots(polarplot, polarannulus, polarfreq, polarperc)
```

Note that the two "roses" aren't shown here and `polarDiff()` and `polarCluster()` are still to be developed, both owing to developments required in `{openair}` itself.

<!-- # The Three Roses -->

<!-- Note that the two "roses" above - wind and pollution - have their own annotation functions to put on their. The `rose_angle` arguments are required if you change the `angle` argument of wind/pollutionrose due to the way `geom_col()` interacts with `coord_polar()` --- simply copy the chosen argument for `angle` and it'll fix any strangeness when `wd` is near 0/North. -->

<!-- ```{r} -->
<!-- gg_windrose(marylebone, facet = "daylight") + -->
<!--   theme_polar(panel_ontop = FALSE) + -->
<!--   annotate_rose_text(y = 15, wd = "SE", size = 2.5) -->
<!-- ``` -->

<!-- The functionality to compare two sets of ws/wd data felt a bit misplaced in `pollutionRose()`, so `{ggopenair}` exports a separate `gg_biasrose()` function. This also lets the plot better reflect what it is actually representing (no more N/S/E/W axis labels). -->

<!-- ```{r} -->
<!-- mydata <- dplyr::mutate(marylebone, -->
<!--                         ws2 = ws + 2 * rnorm(nrow(marylebone)) + 1, -->
<!--                         wd2 = wd + 30 * rnorm(nrow(marylebone)) + 30 -->
<!-- ) -->

<!-- ## need to correct negative wd -->
<!-- id <- which(mydata$wd2 < 0) -->
<!-- mydata$wd2[id] <- mydata$wd2[id] + 360 -->

<!-- ## results show postive bias in wd and ws -->
<!-- gg_biasrose( -->
<!--   mydata, -->
<!--   ws = "ws", -->
<!--   wd = "wd", -->
<!--   ws2 = "ws2", -->
<!--   wd2 = "wd2" -->
<!-- ) + -->
<!--   theme_polar(panel_ontop = FALSE) + -->
<!--   scale_fill_manual(values = c("darkblue", "hotpink")) + -->
<!--   annotate_rose_text(wd = "SW", y = 10, rose_angle = 10) -->

<!-- ``` -->

