---
title: "Trajectory Analysis"
subtitle: "Analyse and visualise HYSPLIT trajectory files."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Trajectory Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggopenair)
library(ggplot2)
```

# Importing Data

Currently, trajectories should be imported using `{openair}` as this functionality hasn't yet been ported to `{ggopenair}`.

```{r}
# import trajectories
traj <- openair::importTraj(site = "london", year = 2010)

# import data for North Kensington
kc1 <- openair::importAURN("kc1", year = 2010)

kc1 <- dplyr::select(kc1, date, nox, no2, o3, pm2.5, pm10)

# now merge with trajectory data by 'date'
traj <- dplyr::left_join(traj, kc1, by = "date")

traj_red <-
  openair::selectByDate(
    traj,
    start = "15/4/2010",
    end = "21/4/2010"
  )
```

# Path Plots

`{ggopenair}` exports `traj_plot()`, which is extremely similar to `trajPlot()`.

```{r}
traj_plot(traj_red, colour = "pm2.5") +
  scale_opencolours_c() +
  theme_classic()
```

# Binned Plots

`traj_level()` is also exported, which again behaves almost identically to `trajLevel()`. All statistic types are supported. Note that using scale transforms on the "fill" aesthetic is almost always recommended.

```{r}
traj_level(traj_red) +
  scale_opencolours_c(
    trans = "sqrt",
    label = scales::label_percent(scale = 1)
  ) +
  labs(fill = "% Trajectories\n(sqrt scale)\n") +
  theme_classic()
```

# Traj Clustering

Unlike `trajCluster()`, `traj_cluster()` only returns clustered data by default.

```{r}
traj_cluster(traj_red)
```

This can be inserted directly into `traj_plot()` to visualise the clusters. 

```{r}
traj_cluster(traj_red)$traj %>%
  traj_plot(colour = "cluster") +
  theme_classic()
```

However, if you prefer the behaviour of `trajCluster()`, you can use the `plot` argument. This requires the most recent version of `openair` from GitHub.

<!-- ```{r} -->
<!-- traj_cluster(traj_red, plot = TRUE) & -->
<!--   theme_classic() & -->
<!--   theme(legend.position = "top") -->
<!-- ``` -->

