---
title: "Time Series and Trends"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time Series and Trends}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggopenair)
library(ggplot2)
library(dplyr)

# not needed - for comparisons
library(openair)
```

# Time Plots

Unlike in `{openair}`, `{ggopenair}` **does not** have a dedicated `time_plot()` function. This is because creating time series in `{ggplot2}` is already so simple. The "legacy" `timePlot()` looks like this:

```{r legacy, fig.cap="The legacy openair time plot"}
timePlot(marylebone, pollutant = "nox", avg.time = "month")
```

The equivalent in `{ggplot2}` is shown below.

```{r newtimeplot, fig.cap="A time plot created using ggplot2."}
plt <- 
  marylebone %>%
  time_average(avg_time = "month") %>%
  ggplot(aes(x = date, y = nox)) +
  geom_line() +
  theme_bw() +
  labs(x = NULL, y = quickText("nox"))
```

While there are more lines of code to achieve the same thing, a `{ggplot2}` plot is much more customisable. `{ggopenair}` even contains tools to help with common issues. For example, we can add limit values using the `scale_y_limitval()` function.

```{r}
plt +
  scale_y_limitval(c(150, 350), "red", c("Getting\nHigher", "Too high!"), c(2,1))
```

There is also no equivalent for `smoothPlot()` as that is already well served by `geom_smooth()`.

```{r}
plt + geom_smooth()
```

# Temporal Variations

`gg_timevariation` works much the same way as `timeVariation()`, but returns a patchwork assemblage. These can be treated similarly to a `{ggplot2}` object, but the ampersand (`&`) symbol is used to style all plots together.

```{r timevar, fig.width=10}
gg_timevariation(marylebone, c("no2", "pm2.5"), normalise = TRUE) &
  theme_bw() &
  scale_fill_discrete(
    aesthetics = c("colour", "fill"),
    labels = scales::label_parse()(c("NO[2]", "PM[2.5]"))
  )
```

The `return` argument can be used to obtain specific panels. This could be useful if not all panels are meaninful for your data, or you want to customise individual panels.

```{r}
hour_panel <- gg_timevariation(marylebone, "o3", return = "hour")

avg = mean(hour_panel$data$Mean)

hour_panel +
  theme_classic() +
  scale_color_manual(values = "hotpink", aesthetics = c("colour", "fill")) +
  scale_y_limitval(marker_values = avg, marker_labels = paste0("Mean (", round(avg, 2), ")")) +
  theme(legend.position = "none")
```

# Time Proportion Plots

`gg_timeprop()` behaves almost identically to `timeProp()`, albeit with fewer arguments as more can be controlled using `labs()`, `scales_*_*()`, and so on.

```{r}
marylebone %>%
  filter(format(date, "%Y") == 2019) %>%
  gg_timeprop(pollutant = "pm10", 
              avg_time = "3 day",
              proportion = "wd") +
  theme_bw() +
  scale_fill_brewer(palette = "Dark2")
```

`proportion` behaves similarly to `facet` (`type` in the original `openair`). If a column isn't present in the data set (or it is numeric), it will use `openair::cutdata()` to parse it and work out how it can be used to cut the data. The recommendation, however, is to get your categories in order *before* you use `proportion` --- then you have full control over and understanding of the output. For example, in the below plot, we use `cut_wd()` to pre-cut the wind directions into bins --- in this case, fewer bins by setting `resolution` to "low".

```{r}
marylebone %>%
  mutate(wd = cut_wd(wd, "low")) %>%
  filter(format(date, "%Y") == 2019,
         !is.na(wd)) %>%
  gg_timeprop(pollutant = "pm10", 
              avg_time = "3 day",
              proportion = "wd") +
  theme_bw() +
  scale_fill_brewer(palette = "Dark2")
```

# Trend Heat Maps

Much like with the time proportion plots, `gg_trendlevel()` is very similar to `trendLevel()`.

```{r}
plt <- 
  gg_trendlevel(marylebone, "nox", "hour", "month", "year") +
  scale_opencolours_c() +
  theme_minimal() +
  labs(fill = quickText("Mean NOx"), y = NULL)

plt
```

A minor exception is that, when possible, axes are automatically parsed as numeric/integer rather than always being factors. This avoids label overlap, and allows users to use continuous scales for further transformations and annotations. For example, here each day is split into four, 6-hour segments.

```{r}
plt + 
  scale_x_continuous(breaks = seq(0, 24, 6)) +
  geom_vline(xintercept = c(6, 12, 18))
```

# Calendar Plots

Calendar plots once again behave similarly, but the annotations are dealt with using bespoke functions. These make it easier to customise or layer the annotations on top of one another.

```{r calendar, fig.width=10}
marylebone |>
  openair::selectByDate(year = 2019) |>
  gg_calendar("nox") +
  annotate_calendar_text("date") +
  annotate_calendar_wd(colour = "black") +
  scale_opencolours_c()
```

An extra feature of `gg_calendar()` is that multiple years can now be easily plotted. Instead of a normal calendar, a year-month matrix is produced.

```{r calendaryears, fig.width=10}
library(scales)
marylebone |>
  gg_calendar("nox") +
  scale_opencolours_c(limits = c(0, 600), oob = squish)
```

# TheilSen

There is currently no version of theilSen in `{ggopenair}`. This is open for discussion, but it will either take the form of an `openair`-like function (e.g., `gg_theilsen()`) *or* a more `{ggplot2}`-esque `geom_theil()`. Until this functionality has been developed, feel free to use `openair::theilSen()`, optionally re-plotting in `{ggplot2}` if desired.
